---
title: "Seminar2b - GGPlot2"
author: "Lisa Leung"
date: '2018-01-10'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Loading tidyverse
```{r library}
library(tidyverse)
```

# The mpg data frame

```{r cars}
mpg
```

# Creating a ggplot

A typical template (in general):

ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <SCALE_FUNCTION> +
  <AXIS_FUNCTION> +
  <FACET_FUNCTION>
  
```{r cars}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

Display vs. Hwy:

1. negative relationship, but doesn't seem to be linear relationship

# Add more factors: class type and cylinders

```{r cars}
mpg$cyl<- factor(mpg$cyl)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class, shape=cyl))
```

# Here is a list of geometries functions
```{r cars}
apropos("^geom_")
```
# Here is a list of statistics functions

```{r cars}
apropos("^stat_")
```
# Here is a list of scale functions

```{r cars}
apropos("^scale_")
```

# Adding a smooth line to the plot:
```{r cars}
ggplot(data = mpg, 
       mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()
```

# Adding colours on a continuous variable:
```{r cars}
ggplot(data = mpg, 
       mapping = aes(x = displ, y = hwy, color = year)) +
  geom_point() +
  geom_smooth()
```

## Making bar charts

# New table for average efficiency based on class
```{r cars}
(averageEfficiency <- 
  mpg %>% group_by(class) %>% summarise(fuel_efficiency = mean(hwy)))
```
# Plotting a geom bar graph to look at the efficiency

Note: need to include  'stat = "identity"' because of an error message. Turns out geom_bar() uses "stat = count" by default. In other words, it would count up the number of rows for a given x value and display that. That basically shows the distribution of some discrete x variable. In which case, the supplied y mapping is irrelevant and hence the error.

But we want geom_bar() to plot bars with heights specified by our y mapping. So what do we do? We can use "stat = identity", which asks geom_bar() to take variable values in the supplied y mapping that correspond to the x values.

```{r cars}
ggplot(averageEfficiency) + 
  geom_bar(aes(x = class, y = fuel_efficiency),
           stat = "identity")
```

# Adding colours and x,y labels to the geom bar graph 
```{r cars}
ggplot(averageEfficiency) + 
  geom_bar(aes(x = class, y = fuel_efficiency, fill = class),
           stat = "identity") +
  ylab("Fuel Efficiency (miles per gallon)") +
  xlab("Vehicle Type")
```
# We can also reverse the y values of the geom bar graph
```{r cars}
ggplot(averageEfficiency) + 
  geom_bar(aes(x = class, y = fuel_efficiency, fill = class),
           stat = "identity") +
  ylab("Fuel Efficiency (miles per gallon)") +
  xlab("Vehicle Type") +
  scale_y_reverse()
```

### THE DELIVERABLE

```{r cars}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv, size=class))
```